---
title: "Data Import"
output: github_document
---

I'm an R Markdown document! 

```{r setup, echo = FALSE,message = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

This document will show how to import data.

# Import excel file

## Import the FAS Litters CSV

```{r}
litters_df = read_csv("data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
# janitor is a package used for cleaning data
# clean_names installed inside the janitor package used to clean the variables name, replace all spaces by _, change every characters into lower case.
```


```{r}
head(litters_df)
tail(litters_df,10)
```

```{r eval=FALSE}
view(litters_df)
```

## Import FAS Pups

```{r}
pups_df = read_csv("data/FAS_pups.csv")

pups_df = janitor::clean_names(pups_df)
```

Use absolute path(the one that use \Users) (not useful cause the path will change every time you move the file)


### Look at read_csv options

```{r}
litters_df = read_csv(file = "data/FAS_litters.csv",
                      col_names = F,
                      skip = 1)
# Always use col_names = F and skip = 1 at the same time, when you have column name but do not wanna show it. Otherwise the col_name will become the first line of data. 
```
### what about missing data

```{r}
litters_df = read_csv(file = "data/FAS_litters.csv",
                      na = c("NA"," ","."))

litters_df = janitor::clean_names(litters_df)

pull(litters_df, gd0_weight) # what does this line do?

```


What if we code `group` as a factpr variable?

```{r}
# make a specific variable to be factor variable rather than default to chr
litters_df = 
  read_csv(file = "data/FAS_litters.csv",
            na = c("NA"," ","."),
           col_types = cols(Group = col_factor()))
```
# Import excel file

```{r}
mlb_df = read_excel("data/mlb11.xlsx",sheet = "mlb11")
# sheet is for excel file that has multiple sheets
# range is used for the range you want to be imported in excel, eg, from A1 to E6
head(mlb_df)
```

# Import SAS data

```{r}
pulse_df = read_sas("data/public_pulse_data.sas7bdat") 

```
Never use read.csv(), also never $. : (





